Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-04-03 draft.tex
Make the complex equation fit the paper
Weak inequality weights

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore, the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
1=w_{1} > w_{2} > \ldots > w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ about the preference of the agent $j$. 
The adversary's goal is to make the score of $y$ as high as possible and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \pprefeq_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \pprefeq_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \pprefeq_j a)\\
y \prefeq_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \pprefeq_j y) ∨ (x \pprefeq_j a)).
\end{split}
\end{align} 
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
\end{claim}
It follows that $\hat{v}_{j}(x)$, the rank of $x$ in the PMR-maximizing linear order of agent $j$, 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\pprefeq_j}(x)}$, where ${\pprefeq_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\pprefeq_j$ (thus including $x$ itself) \commentBN{Is it not better write $\hat{v}_{j}(x) = 1+\card{A}-\card{{\ppref_j}(x)}$? I guess it's for coherence for the $\beta$ expression but it doesn't make much sense pointing out we're removing and adding x.}, and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{{\pinc_j}(x)}$ if $(x \pprefeq_j y)$ and $\beta = 0$ otherwise and with ${\pinc_j}(x)$ designating the alternatives incomparable with $x$ in $\pprefeq_j$ \commentBN{Still not good. In the set of alternatives incomparable with $x$ there could be alternatives strictly better than $y$. It is exactly the problem we had before. Instead, $\beta$ represents the number of alternatives incomparable with $x$ that are no better than $y$: $\beta = \card{A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))}$ }.

@@

Pick any alternative $c$ that is not $a$ or $b$. 
%Let $c_t$ denote the number of times $c$ reaches rank $t$ or better in $P$. Note that $c_m = p+q$ and $c_t ≤ p + q, 1 ≤ t ≤ m$. Define $a_t$ similarly, for the alternative $a$. By construction, $a_r = q$, $a_{r+2} = q+p = a_m$. 
Define $c'_t$ as the number of times $c$ gets rank $t$ or better in $P'$, and define $a'_t$ similarly. Now $a'_1 ≥ \delta, a'_t ≥ 2\delta \text{ for } 2 ≤ t < m, a'_{r+2} = q+p+2\delta = a'_m$. To obtain upper bounds for $c'_t$, assume $c$ is ranked first by the $p+q$ agents voting in profile $P$. By our construction, the  $2 \delta$ new agents of $P'$ give to $c$ the ranks $(3, m)$, or $(4, m-1)$, … Observe that the score of $c$ is maximal for attribution $(3, m)$, by convexity of the weights. Therefore, in the best case for $c$, $c'_1 ≤ p + q, c'_2 ≤ p+q, c'_t ≤ p+q+\delta \text{ for } 3 ≤ t < m, c'_m ≤ p+q+2\delta$. Observe that $a'_t ≥ c'_t, 1 ≤ t ≤ m$ (with strict inequality for $t=2$). Because weights are non increasing, this guarantees, with $\delta = p + q$, that the score of $a$ is not lower than the one of $c$ (it is even strictly higher, thanks to the strict inequality). \commentBN{Here using the strict inequality assumption.}

>>>>

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore, the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
1=w_{1} ≥ w_{2} ≥ \ldots ≥ w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ about the preference of the agent $j$. 
The adversary's goal is to make the score of $y$ as high as possible and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \pprefeq_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \pprefeq_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \pprefeq_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \pprefeq_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
\end{claim}
It follows that $\hat{v}_{j}(x)$, the rank of $x$ in the PMR-maximizing linear order of agent $j$, 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\pprefeq_j}(x)}$, where ${\pprefeq_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\pprefeq_j$ (thus including $x$ itself) \commentBN{Is it not better write $\hat{v}_{j}(x) = 1+\card{A}-\card{{\ppref_j}(x)}$? I guess it's for coherence for the $\beta$ expression but it doesn't make much sense pointing out we're removing and adding x.}, and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{{\pinc_j}(x)}$ if $(x \pprefeq_j y)$ and $\beta = 0$ otherwise and with ${\pinc_j}(x)$ designating the alternatives incomparable with $x$ in $\pprefeq_j$ \commentBN{Still not good. In the set of alternatives incomparable with $x$ there could be alternatives strictly better than $y$. It is exactly the problem we had before. Instead, $\beta$ represents the number of alternatives incomparable with $x$ that are no better than $y$: $\beta = \card{A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))}$ }.

@@

Pick any alternative $c$ that is not $a$ or $b$. 
%Let $c_t$ denote the number of times $c$ reaches rank $t$ or better in $P$. Note that $c_m = p+q$ and $c_t ≤ p + q, 1 ≤ t ≤ m$. Define $a_t$ similarly, for the alternative $a$. By construction, $a_r = q$, $a_{r+2} = q+p = a_m$. 
Define $c'_t$ as the number of times $c$ gets rank $t$ or better in $P'$, and define $a'_t$ similarly. Now $a'_1 ≥ \delta, a'_t ≥ 2\delta \text{ for } 2 ≤ t < m, a'_{r+2} = q+p+2\delta = a'_m$. To obtain upper bounds for $c'_t$, assume $c$ is ranked first by the $p+q$ agents voting in profile $P$. By our construction, the  $2 \delta$ new agents of $P'$ give to $c$ the ranks $(3, m)$, or $(4, m-1)$, … Observe that the score of $c$ is maximal for attribution $(3, m)$, by convexity of the weights. Therefore, in the best case for $c$, $c'_1 ≤ p + q, c'_2 ≤ p+q, c'_t ≤ p+q+\delta \text{ for } 3 ≤ t < m, c'_m ≤ p+q+2\delta$. Observe that $a'_t ≥ c'_t, 1 ≤ t ≤ m$ (with strict inequality for $t=2$). Because weights are non increasing, this guarantees, with $\delta = p + q$, that the score of $a$ is not lower than the one of $c$.
